---
title: "Week3"
author: "your name, D Toher, B Derrick"
date: "insert date"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# installing packages used in the exercise today
```{r, echo=FALSE}
if(!require("ggplot2")){install.packages("ggplot2")}
if(!require("flextable")){install.packages("flextable")}
if(!require("psych")){install.packages("psych")}
if(!require("tableone")){install.packages("tableone")}

library(ggplot2)
library(flextable)
library(psych)
library(tableone)

```

# Introduction to Fisher's Iris dataset

This example is famous and you will find a lot of information about it particularly in R user guides.

The dataset is within R so you do not to read in a data file.

A good starting point to understand the data is to look at the data structure (using the function str) and look at some basic descriptive statistics of each of the variables (using the function summary)


```{r}
str(iris)
summary(iris)
summary(iris$Sepal.Length)
summary(iris$Sepal.Width)
summary(iris$Petal.Length)
summary(iris$Petal.Width)
```


## Example graphics

```{r}
ggplot(data=iris)+
  geom_histogram(mapping=aes(x=Petal.Length),binwidth=.1)


ggplot(data=iris,aes(x=Species,y=Petal.Length))+
  geom_boxplot()
  
  
ggplot(data=iris,aes(x=Petal.Width,y=Petal.Length, color=Species))+
  geom_point()
```

Discussion: 

(1) what do the above graphics show?

(2) which of the above graphics are most appropriate?

(3) what additional graphics may be useful?


# Dewis exercise 1: R_filtering_sampling.

Sometimes you may be asked to work with a subset of data, either excluding a particular group, or taking a sample from a larger dataset.

Insert your code to complete the Dewis exercise here:





## Additional graphics for the iris datsest
Attempt some additional relevant graphics using the iris dataset below:


Interpret your graphics 








# Dewis exercise 2: r_descriptive_statistics

In this exercise you will read in a dataset from Dewis using the function read.csv 

You will then summarise the data and see if this reveals any issues.

Use this space to experiment with the R code provided to complete this task.




# Advanced: Nice looking tables using the additional packages installed

This shows how to start to get some really nice looking tables for the iris dataset

```{r}
describeBy(iris[,-5],group=iris[,5])
```

```{r}
describeBy(iris[,-5],group=iris[,5],skew=FALSE)
IrisSummary<-describeBy(iris[,-5],group=iris[,5],skew=FALSE)
IrisSummary2<-describeBy(iris[,-5],group=iris[,5],skew=FALSE,mat=TRUE)
#View(IrisSummary2)
dim(IrisSummary)
```


```{r}
#colnames(iris)[-5]

# which columns are item and vars?
Ignore<-colnames(IrisSummary2)%in% c("item","vars")

#Ignore
IrisSum<-IrisSummary2[,!Ignore]

#colnames(IrisSum)
colnames(IrisSum)[colnames(IrisSum)=="group1"]<-"species"




flextable(IrisSum)
```


```{r}

# need to add in a leading column with the variable information
colnames(iris)[-5]
niceNames<-c("Sepal length","Sepal width","Petal length","Petal width")

#rep(niceNames,each=3)
IrisSum2<-cbind(variable=rep(niceNames,each=3),IrisSum)
#IrisSum2

flextable(IrisSum2)
```


```{r}
merge_v(flextable(IrisSum2))
```


```{r}
merge_v(flextable(IrisSum2),j=1) # only merge 1st column!
```


```{r}
theme_booktabs(merge_v(flextable(IrisSum2),j=1))
```


```{r}
CreateTableOne(vars=names(iris)[-5],
               strata="Species",  
               data=iris,
              test=F)
```


# End matter - Session Information

It is usually good to make sure that you compile your document so that the information about your session (including software / package versions) is stated.

```{r}
sessionInfo()
```
