---
title: "Introduction to R markdown"
author: '20015367'
date: '07/10/2025'
output:
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) # do not change this option, this shows the code in your knitted document
```

# Firstly check you have any packages required in the session, the only packaged required in this session is ggplot2.
# Run the following code, which double-checks that you have the packages installed,
# and loads them into the R session.

```{r}
if(!require(ggplot2)){install.packages("ggplot2")}
library(ggplot2)
```

### Task 1 Introduction to R markdown ###

This is an R markdown document. Firstly, knit this document to Word, see what happens!


For this exercise we will use data embedded within R, mtcars, start by exploring the structure of the dataset and the variables


```{r cars}
str(mtcars)
```

```{r}
names(mtcars)
```

In this workshop we will concentrate on using mpg (miles per gallon); cyl (number of cylinders) and hp (horsepower).

First, find out more about the dataset using the help file, e.g. with the command **?mtcars**

Next, look at the first 6 rows of data using the command **head**

```{r}
head(mtcars)
```

# Examples of embedding a calculated value into sentences:

The mean mpg for cars in this dataset is `r mean(mtcars$mpg)`.

Edit the statement so that it only prints to 2 decimal places:

The mean mpg for cars in this dataset is `r round(mean(mtcars$mpg),2)`.


# Question. What other summary statistics may you wish to report within the text of a document?


# Example graphics 

Creating a boxplot of miles per gallaon (using geom_boxplot)
```{r}

ggplot(mtcars, aes(x = "", y = mpg)) +
  geom_boxplot(fill = "skyblue", color = "blue") +
  labs(title = "Boxplot of MPG from mtcars Dataset",
       y = "Miles per Gallon (mpg)") +
  theme_minimal()
```

In this code:

- ggplot(mtcars, aes(x = "", y = mpg)): This sets up the basic ggplot object, specifying the dataset (mtcars) and aesthetics (mapping the empty string "" to the x-axis and the 'mpg' variable to the y-axis).

- geom_boxplot(fill = "skyblue", color = "blue"): This adds the boxplot layer to the ggplot object, specifying the fill and border color.

- labs(title = "Boxplot of MPG from mtcars Dataset", y = "Miles per Gallon (mpg)"): This sets the plot title and y-axis label.

- theme_minimal(): This applies a minimal theme to the plot for better aesthetics.

Creating a histogram (using geom_histogram), of miles per gallon and number of cars:
```{r}

ggplot(mtcars,aes(mpg))+geom_histogram(binwidth = 4)+xlab("Miles per Gallon")+ylab("Number of Cars")+ggtitle("Distribution of Cars by Mileage")
```

Editing this to show the histogram for the number of cylinders (edit the binwidth to be equal to 1).
```{r}
ggplot(mtcars,aes(cyl))+geom_histogram(binwidth = 1)+xlab("Cylinders")+ylab("Number of Cars")+ggtitle("Distribution of Cars by Cylinders")
```



### Task 2. Dewis ###

This is the return of the boxplot and histrogram Dewis exercise, but now we include the data for you to do some analyses within R.

As per instructions in Dewis, can you create a new R markdown file, and knit a word document with your answers?
```{r}
ggplot(data=mtcars,aes(x=mpg)) + geom_boxplot()
```
### Task 3. Reproducible research ###

List 'good practice' or standards of reproducible research that you have evidenced in this document (and/or in your R markdown file for the Dewis question).


What other 'good practice' or standards of reproducible research might be appropriate?

# Question: When might you use the 'echo = false' command?




### Task 4. Data Analyses plan ###

Consider the potential research questions that could arise with the mtcars dataset.

Prepare a data analyses plan to address these research questions. 

Open a new R markdown file and prepare a polished report with the analyses as per your statistical analyses plan.




